Information Set Generation in Kriegspiel

### 2012-09-07 paper/paper.tex
Adding all pending

\documentclass[11pt]{article}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage{fullpage}
\usepackage{url,hyperref}
\usepackage{subfig}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{listings}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}
%\usepackage{ijcai09}

@@

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible state of the world is crucial to the
performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
We discuss some variations in the search strategy and their effect on performance.  \end{abstract}

@@

The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
or root node.  Thus, a game tree node implicitly encodes not only the current state of the game, but also all the
history of all decisions made by all players up to that point in the game.  Knowing one's information set means knowing
all possible game histories.

For many games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
opponents is easy to conceive and enumerate.

In the game of kriegspiel, information set generation is non-trivial.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game
tree, although it is arguably a necessary subroutine for quality game play in imperfect information games. The standard
game tree problem is to find an equilibrium (i.e., minimax) strategy for the game. This can be done using alpha-beta
pruning for perfect information games or linear programming for imperfect information games. Information set generation,
by contrast, is the problem of finding sequences of moves in the game tree that are consistent with a player's
observations.

@@

Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
others.  One player controls the white pieces; the other controls the black pieces.  The referee keeps track of both
sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces however he wishes on
his own board (if at all).  The players alternate moves as in a regular chess game, but instead of announcing their
moves out loud, they write the source and destination square of their desired play son a piece of paper and pass them to
the referee, so that the move is not heard by the other player. The referee, who knows the location of all pieces,
checks the legality of the move.  If the move is legal, the referee makes the move on his own board and announces that a
legal move has been made and that it is the other player's turn to move.  If the move is illegal because it is blocked
by an opponent's piece or would place or leave the active player in check, the referee announces, "No."  There is no
penalty for attempting an illegal move.  The player continues to attempt moves until one finally succeeds.  If a player
has no legal moves (because of stalemate or checkmate), the referee announces the game's result.  With the advent of
computers, it is convenient to play the game over a network, with a computer acting as the referee.

@@

The concept of information sets is perhaps best explained through an example.  Figures~\ref{listing} and~\ref{sequence}
show a full listing and the associated pictorial simulation of an actual kriegspiel game from ~\cite{li94chess}.  We
have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves are numbered in pairs: a move by white on
one line followed by black's move on the next line.  The actual move is shown first followed by a list of referee
announcements.  Captures announcements are prefixed with an X and give the location of the captured piece.  Similarly,
pawn tries announcements are prefixed with T.  Attempted moves that were ultimately declared illegal are shown in a list
following the ':'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}), black attempts to block a potential
threat along the diagonal by moving a pawn to f6, then attempts to escape or capture a threat at f7.  When both of these
attempts fail, black can infer that his king is being threatened by a protected bishop or queen at f7.  The attempt to
move to e7 is black's way of finding out whether the threatening piece is a queen or bishop.  When this move fails,
black knows it is white's queen and ultimately retreats to d8.  Note that the failed moves by white on turns 3 and 4 are
a result of being blocked by the pawn at e2.

Note that the listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  Figure~\ref{abbrevoutput} shows the information set for black
after both player's have made five moves.  From this list of possibilities, black can infer that white must have a
bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with black's observations
up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for each move in the game
and the amount of time to find the full set on a single processor.  Appendix B shows the full output of the program for
the same point in the game (five moves for each player) and illustrates all of the possible positions in the belief
state.

@@

Figure~\ref{3schemes}
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
\caption{Experiments using different load balancing strategies.}
\label{3schemes}
\end{figure}

@@

Figure~\ref{3depths}
\begin{figure}
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width=\textwidth]{plots/3depths.pdf}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\caption{3depths}
\label{3depths}
\end{figure}

@@

Figure~\ref{dep8}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{plots/dep8.pdf}
\caption{dep8}
\label{Grainsize control data for depth 8 problems.}
\end{figure}

Figure~\ref{dep10}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{plots/dep10.pdf}
\caption{dep10}
\label{Grainsize control data for depth 10 problems.}
\end{figure}

Figure~\ref{dep12}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{plots/dep12.pdf}
\caption{dep12}
\label{Grainsize control data for depth 12 problems.}
\end{figure}

@@

\section{Related Work}
An overview of the theory of imperfect information games and other basic topics in game theory can be found in
~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
equilibrium can be found in time that is polynomial in the size of the game tree.  And in fact, they found that it took
about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is feasible only
for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently, computer
systems tend to utilize the concept of infomration sets in some form or another.  A common strategy is to estimate the
value of a move by averaging the estimated value of the associated descendant for each node in the current information
set (or a random sample from the set).  While this sort of ``averaging over clairvoyance'' {\em can} lead to very poor
decisions, it is nevertheless a reasonable strategy in some games (e.g., Scrabble~\cite{sheppard02world}).

In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large
game trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
can indeed be expensive to produce the full information set.  For the experiments shown in Table XX, generating the full
information set took over 25 minutes on a single processor.  However, by utilizing the Charm++ search engine on a
parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note that this set had
more than 4 million members.)  Our results suggest that the problem is highly scalable and our approach might therefore
be preferred.

@@

First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
possible worlds in the current state without respect to the overall history mof moves).  However, it has been shown that
significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
opponent, based on the knowledge that the opponent would have had at prior positions in the move
hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
kind that we have described here.\footnote{This work is underway.}

Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
to have some games that go on for 40 or 50 moves and have positions where the information sets are on the order of a few
hundred thousand or a few million in size.

The information set generation algorithm that we have implemented here is one the most straightforward one and is
alluded to in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy
for the Hamiltonian Circuit problem.  An alternative information set generation algorithm is described in
~\cite{richards09information} and could potentially improve the search efficiency greatly by exploiting variable
ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we referred to as ``algorithm 2'' in
class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the root node going forward, this
algorithm would seek opportunities for pruning based on the propagation of logical consequences from each move both
forward and backward in time.  This, incidentally, is more consistent with the way human players would analyze the game.
For example, if a player's proposed bishop move is rejected and the bishop is sufficiently removed from its own king,
then the player may infer that one of his opponent's pieces is along that rejected path.  At least one of those spaces
must have been the destination of a prior move by the opponent.  If a recent move by that bishop crossed the same path,
then that would be hard evidence that those squares were empty at the time.  And therefore, the arrival of the
opponent's piece must have been between the previous (successful) bishop move and the current (unsuccessful) bishop
move.  To implement this algorithm, we envision the use of planning graph type data structures ~\cite{blum97fast}.
Planning graphs have a sequence of alternating levels of state constraints and action constraints.  State constraints
would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.  Action constraints would be of the form
$move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would require a data structure with a bit for each
possible action/state.  Propagating the constraints would require much of the same machinery used in STRIPS-like
planning algorithms~\cite{chen05solving}, and one of the many knobs that would need to be tuned would be the depth of the
propagation for each constraint.  (Should the consequences of a known action or non-action be propagated one move into
the future and one move into the past?  More?  Less?)  Additionally, there would be a significant and interesting
tradeoff in the effectiveness of pruning and the and the size of the data structures that would have to be encoded in
each chare.  The best algorithm might be a hybrid strategy that does forward propagation from the root at the parallel
level and constraint propagation at the sequential level.\footnote{This work is also in progress.}

@@

@@

\section{Appendix B}
\begin{verbatim}
1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6

@@

\bibliographystyle{acm}
\bibliography{paper}

>>>>

%\documentclass[11pt]{article}
\documentclass[conference]{IEEEtran}

\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage{fullpage}
\usepackage{url,hyperref}
\usepackage{subfig}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{listings}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}
%\usepackage{ijcai09}

@@

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is crucial to the
performance of game playing agents.  In this work, we discuss the problem of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
Our results suggest that our algorithm can be efficiently parallelized.  \end{abstract}

@@

The term {\em belief state} is sometimes used interchangeably with information set to refer to a probability
distribution over possible worlds.  The latter term comes from the game theory community and is preferred here.  A node
in a game tree denotes not only the current state of the game, but also uniquely defines a path from the initial state
or root node.  Thus, a game tree node implicitly encodes not only the current state of the game but also the complete
history of all decisions made by all players up to that point in the game, including the outcome of any chance moves
such as dice rolls or card shuffling.  Knowing one's information set means knowing all possible game histories.

For many specific games, solving the information set generation problem is trivial.  For example, in a poker game, it is easy to
see that the unseen cards held by a player's opponents may be any permutation of the cards not seen by that player
(i.e., hole cards and any revealed community cards).  After the betting rounds, it would not be reasonable to assume
that each of these permutations of unseen cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  But the set of {\em possible} hands for all of the
opponents is easy to conceive and enumerate.

In the game of kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
information sets for the game of kriegspiel.

@@

Kriegspiel was invented by Henry Michael Temple in 1899.  Historically, the game has required three parties: two
competing players and one referee.  Each player sits with a chessboard that is partitioned off so as not be seen by the
other.  One player controls the white pieces; the other controls the black pieces.  The referee has a board on which she
keeps track of both sets of pieces.  Each player has a full set of chess pieces.  He may position the opponent's pieces
however he wishes on his own board (if at all).  The players alternate moves as in a regular chess game, but instead of
announcing their moves out loud, they write the source and destination square of their desired plays on a piece of paper
and pass them to the referee, so that the requested move is not known to the other player. The referee, who knows the location of
all pieces, checks the legality of the move.  If the move is legal, the referee makes the move on his own board and
announces that a legal move has been made and that it is the other player's turn to move.  If the move is illegal
because it is blocked by an opponent's piece or would place or leave the active player in check, the referee announces,
``No.''  There is no penalty for attempting an illegal move.  The player continues to attempt moves until one finally
succeeds.  If a player has no legal moves (because of stalemate or checkmate), the referee announces the game's result.
With the advent of computers, it is convenient to play the game over a network, with a computer acting as the referee.

@@

The concept of information sets is perhaps best explained through an example.  (See also Appendix A.)
Figures~\ref{listing} and~\ref{sequence} show a full listing and the associated pictorial progression of an actual
kriegspiel game from ~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  The moves
are numbered in pairs: a move by white on one line followed by black's move on the next line.  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

The listing in Figure~\ref{listing} corresponds to the referee's view of the game and would be useful for
post-game analysis or commentary.  During the game, players would not have access to the full information.  The
transcript from white's perspective is shown in Figure~\ref{filteredlisting}.  Note that black's actual moves have been
replaced with ``??'' and lists of black's illegal moves have been replaced with the {\em number} of illegal attempted
moves.  For example, on move 9, black attempted three illegal moves.
\begin{figure}
\begin{verbatim}
1. d4 {(:)}
   ?? {(:0)}
2. Bg5 {(:)}
   ?? {(:0)}
3. Nc3 {(:Bd8)}
   ??  {(:0)}
4. d5 {(:Bd8;Tc5)}
   ??  {(:0)}
5. d6 {(:)}
   ?? {(:0,Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   ?? {(Xd6:Td6)}
8. Qd5 {(:)}
   ?? {(:0)}
9. Qf7+ {(:)} 
   ?? {:0}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\caption{Listing of a kriegspiel game resulting in checkmate by white after 10 moves.}
\label{filteredlisting}
\end{figure}

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for black is in black's information set.  Figure~\ref{abbrevoutput} shows the
information set for black after both player's have made five moves.  From this list of possibilities, black can infer
that white must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
black's observations up to that point. Table~\ref{bothtimes} shows the size of the information sets for each player for
each move in the game and the amount of time to find the full set on a single processor.  Appendix B shows the full
output of the program for the same point in the game (five moves for each player) and illustrates all of the possible
positions in the belief state.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{plots/3schemes.pdf}
\caption{Experiments using different load balancing strategies.}
\label{3schemes}
\end{figure}

@@

\begin{figure}[h]
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width=\textwidth]{plots/3depths.pdf}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width=\textwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\caption{(Left) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Right) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
\label{3depths}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{plots/dep8.pdf}
\caption{Grainsize control data for depth 8 problems.}
\label{dep8}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{plots/dep10.pdf}
\caption{Grainsize control data for depth 10 problems.}
\label{dep10}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{plots/dep12.pdf}
\caption{Grainsize control data for depth 12 problems.}
\label{dep12}
\end{figure}

@@

\section{Related Work}
An overview of the theory of imperfect information games and other basic topics in game theory can be found
in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Algorithms for ``solving'' imperfect information games---in the sense of finding a
Nash equilibrium---can be found in~\cite{koller94fast} and~\cite{koller97representations}.  Koller \etal, showed that an
equilibrium can be found in time that is polynomial in the number of nodes in the game tree.  And in fact, they found
that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
computer systems tend to utilize the concept of infomration sets or belief states in some form or another.  A common
strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
Scrabble~\cite{sheppard02world}).

In this work, we have argued for an information set generation approach to kriegspiel, but other authors have approached
the problem differently.  Parker \etal, consider the problem of sampling from belief states in a variety of large game
trees, including kriegspiel~\cite{parker05game}.  They show that some performance gains can be achieved by only
approximate sampling from the belief state.  Instead of generating the full information set, or even sampling from it,
they sample from possible positions for the current state (i.e., sampling from the belief state) by matching only the
most recent observations.  This motivation for this approach seems to be that it would not be feasible to sample from
the full information set, because of the computational expense involved.  And indeed, our experiments bear this out-- it
can indeed be expensive to produce the full information set.  For the experiments shown in Table~\ref{speedups},
generating the full information set took over 25 minutes on a single processor.  However, by utilizing the Charm++
search engine on a parallel machine (1024 processors), we were able to generate the full set in just 2.9 seconds.  (Note
that this set had more than 4 million members.)  Our results suggest that the problem is highly scalable and our
approach might therefore be preferred.

@@

First and foremost, it would be interesting to see how our information set generation algorithm would perform as a
subroutine to an actual kriegspiel playing program.  The utility of generating an information set is ultimately
determined by how well that information improves a player's decision-making capabilities.  We noted earlier that a
common use of information sets is to estimate the value of a move by averaging the estimated value of the associated
descendant for each node in the current information set (or a random sample from the set).  This kind of strategy can be
applied naively using a belief state sampling algorithm (i.e., an algorithm that samples from the distribution of
possible worlds in the current state without respect to the overall history of moves).  However, it has been shown that
significant improvement in play can be achieved by estimating the value of future moves from the perspective of the
opponent, based on the knowledge that the opponent would have had at prior positions in the move
hisotry~\cite{richards07opponent}.  Such a strategy would require the use of an information set generation algorithm of the
kind that we have described here.\footnote{This work is underway.}

Ideally, we would run our algorithm on some deeper instances from real kriegspiel games.  We are not aware of a
repository of such games.  Unfortunately, notation for kriegspiel is not standardized and there are several minor
variations in rules (i.e., with respect to the nature of referee announcements) that are unlikely to significantly to
impact the computability and scalability results but which nevertheless make it difficult to write an input reader that
converts the transcript of a game into a sequence of moves readable by our algorithm.  An alternative would be to play
some games ourselves.  Currently, the manual encoding of games is not difficult but is tedious.  Still, it would be nice
to have some games that go on for 40 or 50 moves and have positions where the information sets are of moderate size (on
the order of a few hundred thousand to a few million).

The information set generation algorithm that we have implemented here is the most straightforward one and is alluded to
in~\cite{parker05game} and~\cite{russell05efficient}.  It is analogous to the ``expand-at-the-tip'' strategy for the
Hamiltonian Circuit problem.  An alternative information set generation algorithm is described
in~\cite{richards09information} (but not implemented in parallel) and could potentially improve the search efficiency
greatly by exploiting variable ordering.  This would be analgous to the disconnected edge-pairs\footnote{This is what we
referred to as ``algorithm 2'' in class.} heuristic for the Hamiltonian Circuit problem.  Rather than search from the
root node going forward, this algorithm would seek opportunities for pruning based on the propagation of logical
consequences from each move both forward and backward in time.  This, incidentally, is more consistent with the way
human players would analyze the game.  For example, if a player's proposed bishop move is rejected and the bishop is
sufficiently removed from its own king, then the player may infer that one of his opponent's pieces is along that
rejected path.  At least one of those spaces must have been the destination of a prior move by the opponent.  If a
recent move by that bishop crossed the same path, then that would be hard evidence that those squares were empty at the
time.  And therefore, the arrival of the opponent's piece must have been between the previous (successful) bishop move
and the current (unsuccessful) bishop move.  To implement this algorithm, we envision the use of planning graph type
data structures~\cite{blum97fast}.  Planning graphs have a sequence of alternating levels of state constraints and
action constraints.  State constraints would be of the form $at(bishop,d4) \vee at(bishop,e6)$ or $!occupied(e4)$.
Action constraints would be of the form $move(knight,a4,c5) \vee move(knight,e6)$.  At a minimum, each level would
require a data structure with a bit for each possible action/state.  Propagating the constraints would require much of
the same machinery used in STRIPS-like planning algorithms~\cite{chen05solving}, and one of the many knobs that would
need to be tuned would be the depth of the propagation for each constraint.  (Should the consequences of a known action
or non-action be propagated one move into the future and one move into the past?  More?  Less?)  Additionally, there
would be a significant and interesting tradeoff in the effectiveness of pruning and the size of the data
structures that would have to be encoded in each chare.  The best algorithm might be a hybrid strategy that does forward
propagation from the root at the parallel level and constraint propagation at the sequential level.\footnote{This work
is also in progress.}

@@

\section{Conclusion}
We have shown that information set generation for kriegspiel can be efficiently parallelized.  The opportunities for
parallelism increase as the problem size increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the problem that we are aware of.  Because
the algorithm can be efficiently parallelized, it would be reasonable to explore using it in a full kriegspiel player.
It remains to be seen whether these scalability properties hold over a wide variety of playing styles and problem
instances.  We suspect that fruitful research remains with respect to alternative algorithms that implement variable
ordering heuristics.

@@

\section{Appendix B}
Verbose output from the information set generation algorithm.  The listing shows possible sequences of moves that are
consistent with the black's observations after ten plies.  The resulting position after each consistent sequence of
moves is shown below.  Based on these results, black can know with certainty the location of white's pawn at d6 and
white's bishop at g5.

\begin{verbatim}
1. Pa2:a4 Pa7:a5 2. Pd2:d4 Pb7:b6 3. Bc1:g5 Pc7:c5 4. Pd4:d5 Nb8:a6 5. Pd5:d6 Pf7:f6

@@

%\bibliographystyle{acm}
\bibliographystyle{IEEEtran}

\bibliography{paper}
